Expanding the scope of experimental archaeology using the Perception-Process-Product conceptual framework

### 2023-07-19 manuscript/manuscript.Rmd
added a few references.

On the other hand, actualistic experiments pay more attention to how experimental insights can be generalized to archaeological samples by incorporating authentic materials and plausible social settings with a certain degree of compromised control [@outram2008]. Back to the cases of cultural transmission experiments, an actualistic experiment would involve the use of natural rocks with varied morphology instead of standardized artificial materials as well as human demonstrators instead of videos of knapping instruction, despite the fact that the latter will remain consistent across individuals. Interestingly, researchers in cognitive psychology and neuroscience [@shamay-tsoory2019; @sonkusare2019], a field known for its development of rigorous and well-controlled experimental methods, have started to use naturalistic stimuli more frequently and advocate a paradigm shift to semi-controlled experiment. Unlike controlled experiments, variability could be easily observed in actualistic experiments by design. This feature is crucial and cannot be simply replaced by ethnographic records, because many paleolithic technological components are not displayed in contemporary non-industrial societies, which usually feature technological systems with groundstone artifacts as the target products [e.g., @stout2002; @arthur2018]. Furthermore, statistical techniques for developing causal inference from observational data, which essentially represent the nature of results from actualistic experiments, have also been greatly boosted in recent years [@cunningham2021; @hernan2023]. Lastly, actualistic experiment can serve as a heuristic for hypothesis generation, aligning with the perspective of Lin et al. [-@lin2018: 680-681] the argument put forth by Ingersoll and MacDonald [-@ingersoll1977], who proposed that the interaction between actualistic and controlled experiment "operates in a cyclical form of induction and deduction."

>>>>

On the other hand, actualistic experiments pay more attention to how experimental insights can be generalized to archaeological samples by incorporating authentic materials and plausible social settings with a certain degree of compromised control [@outram2008]. Back to the cases of cultural transmission experiments, an actualistic experiment would involve the use of natural rocks with varied morphology instead of standardized artificial materials as well as human demonstrators instead of videos of knapping instruction, despite the fact that the latter will remain consistent across individuals. Interestingly, researchers in cognitive psychology and neuroscience, a field known for its development of rigorous and well-controlled experimental methods, have started to use naturalistic stimuli more frequently and advocate a paradigm shift to semi-controlled experiment [@shamay-tsoory2019; @sonkusare2019; @nastase2020; @yarkoni2022]. Unlike controlled experiments, variability could be easily observed in actualistic experiments by design. This feature is crucial and cannot be simply replaced by ethnographic records, because many paleolithic technological components are not displayed in contemporary non-industrial societies, which usually feature technological systems with groundstone artifacts as the target products [e.g., @stout2002; @arthur2018]. Furthermore, statistical techniques for developing causal inference from observational data, which essentially represent the nature of results from actualistic experiments, have also been greatly boosted in recent years [@cunningham2021; @hernan2023]. Lastly, actualistic experiment can serve as a heuristic for hypothesis generation, aligning with the perspective of Lin et al. [-@lin2018: 680-681] the argument put forth by Ingersoll and MacDonald [-@ingersoll1977], who proposed that the interaction between actualistic and controlled experiment "operates in a cyclical form of induction and deduction."
